---
title: "GLM"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```


```{r}

oring=read.csv('C:/Data/oringv1.csv')

```

```{r}
oring
```


```{r}
ggplot(data=oring, aes(x=temp, y=td))+geom_point(data=oring, aes(col=factor(td),size=.5 ))+scale_color_manual(values=c('red','lightblue')) + labs(col='Failure')
```

Applying glm to Challenger O-ring test data

```{r}
oring.glm=glm(formula = oring$td ~ oring$temp, family = "binomial")
summary(oring.glm)
```

Another way to formulate the model

```{r}
oring.glm=glm(td ~ temp, family = "binomial", data=oring)
summary(oring.glm)
```

The Null deviance - the deviance   - approximates F test of goodness of fit.

```{r}
G=oring.glm$null.deviance-oring.glm$deviance
G
1-pchisq(G,1)
```


Plot the residuals

```{r}
plot(oring.glm)
```

Plot the data and fitted model

```{r}
p=exp(15.0429-0.2322*oring$temp)/(1+exp(15.0429-0.2322*oring$temp))

plot(c(50,80), c(0,1), type="n", xlab="temperature", ylab="probability",main="ORing Failure")
points(oring$temp,oring$td, pch=16)
lines(oring$temp,p, col=4, lwd=2)

```

inverse logit transforms log odds to probability for temperature = 60

```{r}
p=exp(15.0429-0.2322*60)/(1+exp(15.0429-0.2322*60))
p
```

Does a higher order perform better?

```{r}
mod.2=glm(td ~ temp + I(temp^2) ,family = "binomial", data=oring)
summary(mod.2)
```

```{r}
anova(oring.glm,mod.2, test='Chisq')
```



Blue Oak data from Plant text

```{r include=FALSE}
library(sf)

```


```{r }
data.Set2S <- st_read("C:/Data/set2sierra.shp")
```

```{r}
summary(data.Set2S)
```

```{r}
summary(data.Set2S$QUDO)
```

 Set up a binomial  glm.
 
```{r}

glm.demo <- glm(QUDO ~ Elevation, data = data.Set2S,
   family = binomial)
summary(glm.demo)


```

Check the portion of deviance explained

```{r}

(glm.demo$null.deviance-glm.demo$deviance)/glm.demo$null.deviance
glm.demo$null.deviance-glm.demo$deviance
1-pchisq(1004.718,1)
```

Plot the data and fitted model

```{r}
par(mai = c(1,1,1,1))
plot(data.Set2S$Elevation, data.Set2S$QUDO, # Fig. 8.8
   main = "Sierra Nevada Blue Oaks vs. Elevation",
   cex.main = 1.5, pch=16, col=2,xlab = "Elevation (m)",
   ylab = "Blue Oak Presence/Absence", cex.lab = 1.5)
elev.seq <- data.frame(Elevation = seq(0,2000,50))
lines(seq(0,2000,50), predict(glm.demo, elev.seq,
   type = "response"), lwd=2, col=4)
```

Plot the residuals

```{r}
plot(glm.demo)
```


Try a higher order model

```{r}
glm.demo2 <- glm(QUDO ~ Elevation + I(Elevation^2), data = data.Set2S,
   family = binomial)
summary(glm.demo2)
```

```{r}
glm.demo2$null.deviance-glm.demo2$deviance
1-pchisq(1009.567,2)
(glm.demo2$null.deviance-glm.demo2$deviance)/glm.demo2$null.deviance
```


```{r}
anova(glm.demo, glm.demo2, test='Chisq')
```

```{r}
plot(glm.demo2)
```














