---
title: "GadallaLab8"

output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 
              '.')
```


```{r results='hide', warning=FALSE, message=FALSE}
library(spatstat)
library(splancs)

```

Loading the internal data from packages spatstat and splancs.
```{r}
data(cardiff) 
cardiffnew= as.ppp(cardiff) 
data(gordon)
```

Plotting the two datasets
```{r warning=FALSE, message=FALSE, fig.show="hold", out.width="50%"}
plot(cardiffnew, pch=16)
plot(gordon, pch=16)
```

# 1. Intensity of both point pattern datasets with kernel estimates
```{r warning=FALSE, message=FALSE, fig.show="hold", out.width="50%", results='hold'}
Bw1cardiff = bw.diggle(cardiffnew)
Bw1gordon = bw.diggle(gordon)

Bw1cardiff
Bw1gordon

plot(density(cardiffnew,2.16), main = "Cardiff, sigma 2.16")
plot(density(gordon, 0.7), main = "Gordon, sigma = 0.7")

plot(density(cardiffnew,4), main = "Cardiff, sigma 4")
plot(density(gordon, 1), main = "Gordon, sigma = 1")
```

The output values from the bw.diggle function capture the variation in intensity. I find that increasing the sigma value causes resolution lose and less representative of the actual variation in the dataset.



# 2. Evaluate cardiff data

First order effect exploration
```{r}
quadrat.test(cardiffnew, nx = 3, ny = 3)
```

Results of Chi square test (X^2=32.5 and p value = 0.000) suggest the data set does not follow a process of compelete spatial randomness. 


Data model fitting
```{r}
Fit1=ppm(cardiffnew,  ~1) 
Fit2=ppm(cardiffnew,  ~x+y) 
Fit3=ppm(cardiffnew,  ~ polynom(x, y, 2)) 
```


Model comparisons
```{r warning=FALSE, message=FALSE, results='hold'}
AIC(Fit1)
AIC(Fit2)
AIC(Fit3)
```


AIC comparison suggests that the second order inhomogeneous poisson model (Fit3) works better for Cardiff data set.

Performing anova test to compare between models
```{r warning=FALSE, message=FALSE, results='hold'}
anova(Fit1, Fit2, test = "Chi")
anova(Fit1, Fit3, test = "Chi")
```

Both models are significantly different from CSR model, which confirms that the spatial process is not completely random or following poisson point pattern.


Second order effect Cardiff data

```{r warning=FALSE, message=FALSE, fig.show="hold", out.width="50%", results='hold'}
#Ghat
cardiffnewG.env<-envelope(cardiffnew, fun=Gest, nsim=39, global =TRUE)
cardiffnewG.env
plot(cardiffnewG.env, legend=F, main = "Ghat")

# Lhat
cardiffnewL.env<-envelope(cardiffnew, fun=Lest, nsim=39, global =TRUE)
plot(cardiffnewL.env, .-r ~ r, legend= F, main= "Lhat")
```



Evaluating the two plots of Ghat and Lhat, it shows that the point pattern deviates from poisson pattern, i.e there is point-point dependency. Points tend to cluster.


# 3. Evaluate Gordon data set

first order departure
```{r warning=FALSE, message=FALSE, results='hold'}
quadrat.test(gordon, nx = 3, ny = 3)


Fit1g=ppm(gordon,  ~1) 
Fit2g=ppm(gordon,  ~x+y) 
Fit3g=ppm(gordon,  ~ polynom(x, y, 2)) 

AIC(Fit1)
AIC(Fit2)
AIC(Fit3)

anova(Fit1g, Fit2g, test = "Chi")
anova(Fit1g, Fit3g, test = "Chi")
```

Chi-square quadrat test is significant, meaing Gordon data deviate from complete spatial randomness model.
AIC seems to indicate that the second order inhomogenous model fit the data better than the other two models.
Anova test also that first order inhomogenous model is not significant and that the second order inhomogenous model works better with Gordon data.


Second order effect
```{r warning=FALSE, message=FALSE, fig.show="hold", out.width="50%", results='hold'}
#Ghat
gordon.env <- envelope(gordon, fun=Gest, nsim=39, global =TRUE)
gordon.env
plot(gordon.env, legend=F, main = "Ghat")

# Lhat
gordon.env <- envelope(gordon, fun=Lest, nsim=39, global =TRUE)
plot(gordon.env, .-r ~ r, legend=F, main= "Lhat")
```

I think from the plot, I can say that mostly Gordon data set the points are clustering together, but there are smaller tendency for some points to be independent. 
