---
title: "Space-Time"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Data/')
```


```{r include=FALSE}
library(splancs)
library(spatstat)
library(ggplot2)
library(sf)
library(ade4)
```

Function to compute the Knox test 

```{r}
knoxX=function (s,t, crit_s, crit_t){
  X=sum(s < crit_s & s !=0&t < crit_t & t !=0)
  return(X)} 

knoxsim <- function(t,s,crit_s, crit_t,nsim){
  hold <- matrix(0,nsim)
  for (i in 1:nsim){
    temp <- sample(t, length(t))
    difft <- dist(temp, method= "manhattan")
    hold[i] <- knoxX(s, difft, crit_s, crit_t)}
  return(hold)}

knoxZScoretest <-function(diffs, difft,crit_s, crit_t, n){
  sumspace=sum(diffs < crit_s & diffs != 0 ) 
  sumtime=sum(difft < crit_t & difft !=0) 
  exKnox=(2*sumspace*sumtime)/(n*(n-1))
}

Knoxtest <-function(data, crit_s, crit_t, nsims, alpha){ 
          s<-dist(cbind(data$x, data$y))
          t<-dist((data$t), method = "manhattan")
          times = data$t
          #t<-dist((data$marks), method = "manhattan")
          #times = data$marks
          n=length(times)
          X = knoxX(s,t, crit_s, crit_t)
          simulations<- knoxsim(times,s,crit_s, crit_t,nsims)
          simulations
          hist(simulations ,main="Monte Carlo Distribution of Knox",       col='lightblue',xlab="Simulated Knox values", xlim=c(min(simulations), (X+4)))
          abline(v=X, col=2)
          NGE=sum(simulations>X)
          p_value = round((NGE+1)/(nsims +1),3)
          zt=knoxZScoretest(s, t, crit_s, crit_t,n)
          zvalue=(X-zt)/(sqrt(zt))
          result1=(ifelse(p_value<alpha, "reject CSTR", "Do not reject CSTR"))
          pz=pnorm(zvalue, lower.tail = F)
          result2 =(ifelse(pz<alpha, "reject CSTR", "Do not reject CSTR"))
          output=c("ObsKnox"=X, "NGE"=NGE, "P_value"=p_value,"result1"=result1, "expectedKnox"=round(zt,2),"zvalue"=round(zvalue,2),"zpvalue"=round(pz,3), "result2"=result2)
          return(output)}


```


```{r}
data(burkitt)
```

Locations of cases of Burkitt's lymphoma in the Western Nile district of Uganda 1960-1975. The time variable is recorded as the number of days starting from an origin of 1 Jan 1960. 

```{r}
summary(burkitt)
```

```{r}
ggplot(data=burkitt, aes(x=x,y=y,col=t))+geom_point(size=2)+scale_color_gradient(low='red', high="white") +theme_dark()+labs(col='Time')+ggtitle('Burkitts Lymphoma')
```

Test on burkitts lymphoma e.g. using 180 days and 4 km

```{r}

getknox <- Knoxtest(data=burkitt,2,180, nsims=199, alpha=.05)
getknox
```

 Test Idaho forest fires from 2017 for space time interaction
 
```{r include=FALSE}

fires=read_sf('Firesmallset.shp')
```

```{r}
summary(fires)
```

```{r}
ggplot()+geom_sf(data=fires,aes(col=JULIAN), size=2.8) + scale_color_distiller(palette = "RdBu", direction = 1)+ ggtitle("Fires by Year Day")+ theme_dark()
                 
```

Convert forest fire data to format for Knoxtest

```{r}
xy=st_coordinates(fires)
x=xy[,1]
y=xy[,2]
firesnew=data.frame(x/1000,y/1000,fires$JULIAN)
#firesnew=data.frame(x,y,fires$JULIAN)
colnames(firesnew)=c("x","y","t")

```

```{r}
summary(firesnew)
```

```{r}
sdist<-dist(cbind(firesnew$x,firesnew$y))
mean(sdist)
```


Test fires data for space-time interaction -7000

```{r}
Knoxtest(firesnew,10,10, 199,.05)
```


Mantel Test
The Mantel test can be used to test for spatial autocorrelation.  For autocorrelation the tests uses
spatial distances
attribute distances


It can also be used to test for space-time interaction. For space-time interaction test use
spatial distances
time distances


Set up for Mantel test




```{r}
s<-dist(cbind(burkitt$x,burkitt$y))
t<-dist(burkitt$t,method= "manhattan")

```

Set up of fires data for Mantel test

```{r}
s<-dist(cbind(firesnew$x,firesnew$y))
t<-dist(firesnew$t,method= "manhattan")

```

Compute the Mantel test. This test reports the mantel score - a correlation value and its pvalue based on how extreme the observed  value is relative the set of mantel values generated from some number of permutations: eg. 99

```{r}
plot(mantel.randtest(s,t,99))

```


Convert burkitt data to spatstat ppp file

```{r}
xmin=min(burbdy[,1])
xmax=max(burbdy[,1])
ymin=min(burbdy[,2])
ymax=max(burbdy[,2])
W <- owin(c(xmin,xmax), c(ymin,ymax))
```

```{r}
burk.ppp=as.ppp(burkitt, W=W)
burk.ppp$marks=burkitt$t
```

```{r}
bL=Lest(burk.ppp)
plot(bL, .-r ~r)
```

L function test for spatial clustering

```{r}
burk.lenv = envelope(burk.ppp, Lest, 19, global=T)
plot(burk.lenv, cbind(obs,hi,lo,theo)-theo ~theo, col=c(2,1,1,4),lwd=c(2,1,1,1),lty=c(1,1,1,2),main ="Lymphoma Cases",legend=F)
legend("right" ,legend=c("obs","hi","lo", "theo"), col=c(2,1,1,4), lwd=c(2,1,1,1),lty=c(1,1,1,2), box.lty=0)
```


Space and Time K functions from splancs

Apply Monte carlo test of space-time Kfunction for space-time interaction

```{r}
bur1mc <- stmctest(burpts, burkitt$t, burbdy, c(400, 5800),
  seq(1,40,2), seq(100, 1500, 100), nsim=49, quiet=TRUE, returnSims=TRUE)

```


```{r}
hist(bur1mc$t)
abline(v=bur1mc$t0, col=2)

```


